Capstone Movielens Report

### 2025-05-27 reports/capstone-movielens.site/02.8-0.UMGYDE-model.Rmd
The model mathematical description completed.

If we define $d_{i,j}$ as the _number of days since the earliest record_ in the `edx` dataset for movie $j$ rated by user $i$, then the formula \@res(eq:UMGYE-model) describing the _UMGYE Model_, for the current model, takes the form:

\begin{equation}
Y_{i,j} = \mu + \alpha_i + \beta_j + g_{i,j} + \gamma(v_{i,j}) + s(d_{i,j}) \varepsilon_{i,j}
(\#eq:UMGYDE-model)
\end{equation}

with $s$ a smooth function of $d_{i,j}$

Therefore, the formula \@res(eq:year-effect) for calculation the prediction of a _year effect_ as a residual, for a _smoothed day effect_ 
$$
\hat{f} = s(d_{i,j}) + \varepsilon_{i,j}
$$

@@

\begin{equation}
\hat{s(d_{i,j})} = Y_{i,j} - (\mu + \alpha_i + \beta_j + g_{i,j} + \gamma(v_{i,j}))
(\#eq:smoothed-day-effect)
\end{equation}

@@

We have already explained the idea of _Linear Model Regularization_ in the [UME Model Regularization] section above. We have also seen how the formula \@res(eq:ME-penalty) for adding a penalty to the _UME Model_ is transformed into the formula \@res(eq:GE-penalty) for the _UMGE Model_ and then into the formula \@res(eq:GYE-penalty) for the _UMGYE Model_. For the current model, this formula takes the form:

@@

And the formula \@res(eq:ME-regularized) for calculating the values of the _treatment effect_ that minimizes the equation will take the form:

>>>>

If we define $d_{i,j}$ as the _number of days since the earliest record_ in the `edx` dataset for movie $j$ rated by user $i$, then the formula \@ref(eq:UMGYE-model) describing the _UMGYE Model_, for the current model, takes the form:

\begin{equation}
Y_{i,j} = \mu + \alpha_i + \beta_j + g_{i,j} + \gamma(v_{i,j}) + s(d_{i,j}) + \varepsilon_{i,j}
(\#eq:UMGYDE-model)
\end{equation}

with $s$ a _smoothed day_ function of $d_{i,j}$

Therefore, the formula \@ref(eq:year-effect) for calculation the prediction of a _year effect_ as a residual, for a _smoothed day effect_ 
$$
\hat{s}(d_{i,j}) = s(d_{i,j}) + \varepsilon_{i,j}
$$

@@

\begin{equation}
\hat{s}(d_{i,j}) = Y_{i,j} - (\mu + \alpha_i + \beta_j + g_{i,j} + \gamma(v_{i,j}))
(\#eq:smoothed-day-effect)
\end{equation}

@@

We have already explained the idea of _Linear Model Regularization_ in the [UME Model Regularization] section above. We have also seen how the formula \@ref(eq:ME-penalty) for adding a penalty to the _UME Model_ is transformed into the formula \@ref(eq:GE-penalty) for the _UMGE Model_ and then into the formula \@ref(eq:GYE-penalty) for the _UMGYE Model_. For the current model, this formula takes the form:

@@

And the formula \@ref(eq:ME-regularized) for calculating the values of the _treatment effect_ that minimizes the equation will take the form:
